Augmentation of a Historical Harpsichord Keyboard Replica for Haptic-Enabled Interaction in Museum Exhibitions

### 2025-01-29 src/main.tex
fix #28
fix #29
v1.0.0

%%
%% The "author" command and its associated commands are used to define
%% the authors and their affiliations.
%% Of note is the shared affiliation of the first two authors, and the
%% "authornote" and "authornotemark" commands
%% used to denote shared contribution to the research.
\author{Matthew Hamilton}
% \authornote{Both authors contributed equally to this research.}
\email{matthew.hamilton2@unibo.it}
% \orcid{1234-5678-9012}
\author{Michele Ducceschi}
% \authornotemark[1]
\affiliation{%
  \institution{Università di Bologna}
  \city{Bologna}  
  \country{Italy}
}

\author{Roberto Livi}
\affiliation{%
  \institution{Harpsichord builder}
  \city{Pesaro}
  \country{Italy}
  }

\author{Catalina Vincens}
\affiliation{%
  \institution{Museo di San Colombano}
  \city{Bologna}
  \country{Italy}}

\author{Andrew McPherson}
\affiliation{%
 \institution{Imperial College London}
 \city{London}
 \country{United Kingdom}}

@@

%%
%% The abstract is a short summary of the work to be presented in the
%% article.
\begin{abstract}

%% Version 1
This paper describes the design and creation of an electronically augmented replica of a 16th-century harpsichord keyboard with a typical Renaissance Italian layout to create a digital musical instrument. % [1]
The keyboard was commissioned for exhibition in a musical instrument museum as an enhancement to the visitor experience by providing an interface to digitised versions of instruments within the collection. % [2]
This project also addresses how we can interact with digitally conserve historic musical instruments when the interface is a limiting factor to the verisimilitude of the experience. % [2,3]
This ultimately culminated in a design for an interface capable of new expressions by extending, rather than forfeiting, the characteristics of the original. % [4]

The tactile feedback of the keyboard is created by incorporating the same mechanics as a single manual two-register harpsichord with strings under tension in order to replicate the mechanical resistance characteristic of the original compass. % [1]
An array of optical sensors connected to a microcontroller was installed within the interface to generate continuous streams of data corresponding to the vertical motion of each jack. % [1, ~4]
We outline the details of the keyboard's design principles, technical development, and the motivations behind them. % [1]

%% Version 2
% This paper describes the design and creation of an electronically augmented replica of a 16th-century harpsichord keyboard with a typical Renaissance Italian layout to create a vibrotactile digital musical instrument. % [1]
% An array of optical sensors connected to a microcontroller was installed within the interface to generate continuous streams of data corresponding to the vertical motion of each jack. % [1, ~4]
% The keyboard was commissioned for exhibition in a musical instrument museum as an enhancement to the visitor experience by providing an interface to digitised versions of instruments within the collection.

% This project embodies the entanglement between musical instrument cultural heritage and its relationship with new and increasing accessible technologies. [3, 4]
% Acknowledgment and conscientiousness to older designs and aesthetics have been used in service of, rather than obstruction to, richer interactions with digital musical instruments.[2,3,4]
% This ultimately culminated in a design for an interface capable of new expressions by extending, rather than forfeiting, the characteristics of the original. % [4]
% We outline the details of the keyboard's design principles, technical development, and the motivations behind them. % [1]

@@

\begin{acks}
 We acknowledge the invaluable contributions of
\begin{itemize}
\item
  Università di Bologna
  \begin{itemize}
  \item
    \emph{Alma Labor}:
      Massimiliano Fraulini and the staff of Alma Labor at the Università
      di Bologna for their 3D printing expertise and generosity with
      electronics equipment
  \item
    \emph{Fisica Tecnica}: Maurzio Chendi, Fabrizio Casarini, Giulio Severo and Barbara
      Costantini    
  \end{itemize}
\item
  University of Edinburgh
  \begin{itemize}
  \item
    \emph{uCreate}:    
      Simeon, Stuart, Millie, and Ruth of uCreate at the Univeristy of
      Edinburgh for their support with soldering facilities.    
  \item
    \emph{Edinburgh College of Art}: Richard Collins and the Digital Making Staff at the Edinburgh
      College of Art for 3D printing instruction and CNC support. Digital Development Technician Joe Hathaway    
    \item  
      \emph{St Cecilia's}: Curator Dr. Jenny Nex
  \end{itemize}
\end{itemize}

>>>>

%%
%% The "author" command and its associated commands are used to define
%% the authors and their affiliations.
%% Of note is the shared affiliation of the first two authors, and the
%% "authornote" and "authornotemark" commands
%% used to denote shared contribution to the research.
\author{Matthew Hamilton}
% \authornote{Both authors contributed equally to this research.}
\affiliation{%
  \institution{Università di Bologna}
  \city{Bologna}  
  \country{Italy}
}
\email{matthew.hamilton2@unibo.it}
% \orcid{1234-5678-9012}
\author{Michele Ducceschi}
% \authornotemark[1]
\affiliation{%
  \institution{Università di Bologna}
  \city{Bologna}  
  \country{Italy}
}
\email{michele.ducceschi@unibo.it}

\author{Roberto Livi}
\affiliation{%
  \institution{Master Harpsichord Maker}
  \city{Pesaro}
  \country{Italy}
  }
\email{robertvs.livi@gmail.com}

\author{Catalina Vicens}
\affiliation{%
  \institution{Museo di San Colombano}
  \city{Bologna}
  \country{Italy}}
  \email{catalina.vicens@genusbononiae.it}

\author{Andrew McPherson}
\affiliation{%
 \institution{Imperial College London}
 \city{London}
 \country{United Kingdom}}
 \email{andrew.mcpherson@imperial.ac.uk}

@@

%%
%% The abstract is a short summary of the work to be presented in the
%% article.
\begin{abstract}
% This paper describes the design and creation of an electronically augmented replica of a 16th-century harpsichord keyboard with a typical Renaissance Italian layout to create a digital musical instrument. % [1]
% The tactile feedback of the keyboard is created by incorporating the same mechanics as a single manual two-register harpsichord with strings under tension in order to replicate the mechanical resistance characteristic of the original compass. % [1]
% An optical sensor system was fitted inside the body of the harpsichord, directed towards a gradient surface applied to the body of each jack, generating continuous streams of data corresponding that are used to trigger a sample library. % [1, ~4]
% The keyboard was commissioned for exhibition in a musical instrument museum as an enhancement to the visitor experience by providing means of accessing to instruments that are no longer in playing condition. % [2]
% The details of the keyboard's design principles, technical development, and underlying motivations are described in detail. % [1]
% % The design and mechanics of a harpsichord were replicated in the new keyboard as using modern plastic MIDI controllers has been shown to be an unsatisfying experience for users. % [3]
% % This culminated in a design for a keyboard interface capable of extending the performance techniques of the harpsichord. % [4]

This paper describes the design and creation of an electronically augmented replica of a 16th-century harpsichord keyboard with a typical Renaissance Italian layout to create a digital musical instrument. 
The keyboard was commissioned for exhibition in a musical instrument museum to enhance the visitor experience by providing an interface to digitised versions of instruments within the collection. 
The replica is designed to balance the competing demands of historical authenticity, public accessibility, and preservation. 
It replicates the original instrument's tactile feedback and mechanical resistance by employing historically informed construction techniques. Optical sensors integrated within the mechanism capture the jacks' motion data, enabling MIDI message generation. 
This work situates itself within broader discussions on the role of technology in museums. 
A keyboard interface of this type offers an opportunity to enhance visitor interaction while safeguarding delicate artefacts. 
The paper examines the keyboard's design principles, technical implementation, and implications, emphasising its contribution to public engagement and the long-term preservation of musical heritage.

@@

\begin{acks}
This work received funding from the European Commission's Horizon2020 programme, through grant \anon{ERC-StG 950084 NEMUS}

 We acknowledge the invaluable contributions of
\begin{itemize}
\item
  Università di Bologna
  \begin{itemize}
  \item
    \emph{Alma Labor}:
      Massimiliano Fraulini and the staff of Alma Labor at the Università
      di Bologna for their 3D printing expertise and generosity with
      electronics equipment
  \item
    \emph{Fisica Tecnica}: Maurzio Chendi, Fabrizio Casarini, Giulio Severo and Barbara
      Costantini    
  \end{itemize}
\item
  University of Edinburgh
  \begin{itemize}
  \item
    \emph{uCreate}:    
      Simeon, Stuart, Millie, and Ruth of uCreate at the Univeristy of
      Edinburgh for their support with soldering facilities.    
  \item
    \emph{Edinburgh College of Art}: Richard Collins and the Digital Making Staff at the Edinburgh
      College of Art for 3D printing instruction and CNC support. Digital Development Technician Joe Hathaway    
    \item  
      \emph{St Cecilia's}: Curator Dr. Jenny Nex
  \end{itemize}
\end{itemize}
